Quantum Walks and Application to Quantum Money\\[1em] \normalsize McMaster University\\ Department of Computing and Software

### 2025-04-13 draft.tex
todo list done

\hypersetup{
	unicode = true,
    colorlinks = true,
    citecolor = citecolor,
    filecolor = blue,
    linkcolor = blue,
    urlcolor = blue,
    pdfstartview = {FitH},
}

\title{
  Quantum Walks and Application to Quantum Money\\[1em]
%   \large Supervised by Dr. Jake Doliskani\\[3em]
  \normalsize McMaster University\\
  Department of Computing and Software
}
\author{Ali Mousavi}
\date{April 2025}

@@

This thesis explores the foundations of quantum computation, focusing on quantum walks and their application to quantum money. Quantum walks, particularly continuous-time quantum walks based on group actions, serve as a powerful computational tool with applications in search algorithms and cryptographic protocols. We examine their mathematical structure and their advantages over classical random walks, emphasizing their efficiency in state evolution and probability distribution spreading. As a part of this work, we examine efficient implementations of transforms such as the Quantum Fourier Transform (QFT) and the Quantum Hartley Transform (QHT), analyzing their role in encoding quantum states for secure cryptographic applications. In particular, we discuss a novel instantiation of a quantum money scheme based on QHT, leveraging its unique properties for improved security and efficiency.
To ensure the robustness of this quantum money scheme, we develop a verification mechanism utilizing quantum walks. Unlike previous approaches, which rely on standard quantum state measurements, our method employs continuous-time quantum walks to authenticate quantum money, preventing counterfeiting while maintaining computational feasibility. Additionally, we present a detailed discussion on the efficient implementation of this scheme, including optimized circuit designs and error mitigation strategies.

@@

\medskip\noindent\textbf{This Work – Motivation and Contributions.} 
In this thesis, we adapt Zhandry’s group-action quantum money scheme by replacing its use of the Quantum Fourier Transform with the \emph{Quantum Hartley Transform} (QHT) over finite abelian groups. The motivation behind this substitution is multifold. First, using the Hartley transform causes the banknotes to have \emph{real} amplitudes instead of complex amplitudes. We believe this shift from complex to real quantum states may offer both computational and theoretical advantages. For example, certain mathematical identities hold for any two real orthonormal bases that do not generally hold for complex bases; such differences can affect the properties of quantum states and were a barrier in prior analyses (indeed, an attempted “quantum lightning” construction in \cite{Zhandry2022} failed due to properties of complex phases that do not arise with real amplitudes). Beyond this specific context, our work is a first step toward employing real-valued quantum transforms in quantum cryptography and quantum algorithms. To the best of our knowledge, this is the first significant quantum cryptographic construction that makes essential use of the quantum Hartley transform. We hope that exploring real-amplitude quantum states will inspire further research into new quantum algorithms and optimizations for real-valued transforms.

@@

Quantum computation is a model of information processing rooted in the principles of quantum mechanics. Unlike classical systems, which encode and manipulate information using deterministic bits, quantum systems store information in quantum states—vectors in complex Hilbert spaces—that can exhibit phenomena with no classical analogue. The most notable of these are \emph{superposition}, which allows quantum systems to be in multiple configurations simultaneously, and \emph{entanglement}, which enables correlations between parts of a system that cannot be explained classically. These characteristics endow quantum computers with computational abilities that can outperform classical counterparts in certain tasks.

@@

  \item \textbf{Composition of Systems.} If two quantum systems $A$ and $B$ are described by Hilbert spaces $\mathcal{H}_A$ and $\mathcal{H}_B$, respectively, then the combined system is described by the tensor product $\mathcal{H}_A \otimes \mathcal{H}_B$. A product state of the form $\ket{\psi_A} \otimes \ket{\psi_B}$ represents a system where $A$ and $B$ are uncorrelated. In general, however, the state may be entangled, and cannot be written as a product of subsystems.
\end{enumerate}

@@

\section{The Quantum Fourier Transform}
Let $G$ be an abelian group. The group of characters of $G$, denoted $\hat{G}$, consists of homomorphisms of the form $\chi(a, \cdot): G \to \mathbb{C}$ for each $a \in G$. If the group has the structure $G \cong \mathbb{Z}_{N_1} \oplus \cdots \oplus \mathbb{Z}_{N_k}$, then the character function $\chi(a, \cdot)$ can be expressed explicitly as
\[
\chi(a, x) = \omega_{N_1}^{a_1x_1} \cdots \omega_{N_k}^{a_kx_k}
\]
where $\omega_M = \exp(2\pi i / M)$ is a primitive $M$-th root of unity. Given a function $f: G \to \mathbb{C}$, its Fourier transform is defined as
\[
\hat{f}(a) = \frac{1}{\sqrt{\abs{G}}} \sum_{x \in G} \chi(a, x) f(x).
\]
Correspondingly, the quantum Fourier transform of a normalized state $\sum_{g \in G} f(g) \ket{g}$ is given by $\sum_{x \in G} \hat{f}(x) \ket{x}$.

@@

Another concept we use is the \textbf{quantum Hartley transform} (QHT), which replaces complex exponentials with real sinusoidal kernels.
Let $N$ be a positive integer, and let $\Z_N$ denote the additive cyclic group of integers modulo $N$. The Hartley transform of a function $f: \Z_N \to \R$ is defined as the function $\cht_N(f): \Z_N \to \R$ given by
\[
\cht_N(f)(a) = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big(\frac{2 \pi ay}{N}\Big) f(y),
\]
where $\cas(x) = \cos(x) + \sin(x)$. Similar to the Fourier transform, $\cht_N$ is a linear operation, and one can verify that it is unitary. This definition extends naturally to general abelian groups. Suppose $G = \Z_{N_1} \oplus \Z_{N_2} \oplus \cdots \oplus \Z_{N_k}$ is a decomposition of an abelian group into cyclic components. For a function $f: G \to \R$, the Hartley transform $\cht_G(f): G \to \R$ is defined by
\begin{equation}
    \label{eq:cht-add}
    \cht_G(f)(a_1, \dots, a_k) = \frac{1}{\sqrt{\abs{G}}} \sum_{y_1 = 0}^{N_1 - 1} \cdots \sum_{y_k = 0}^{N_k - 1} \cas\Big(\frac{2 \pi a_1y_1}{N_1} + \cdots + \frac{2 \pi a_ky_k}{N_k}\Big) f(y_1, \dots, y_k),
\end{equation}
with $(a_1, \dots, a_k) \in G$. Introducing the notation $\alpha(\bm{y}) = y_1 / N_1 + \cdots + y_k / N_k$, a more concise form for $\cht_G$ is
\[
\cht_G(f)(\bm{a}) = \frac{1}{\sqrt{\abs{G}}} \sum_{\bm{y} \in G} \cas(2\pi \lrang{\bm{a}, \alpha(\bm{y})}) f(\bm{y}).
\]

@@

The quantum Hartley transform over an abelian group $G$ is given by
\begin{equation}
    \label{eq:qht}
    \qht_G: \sum_{\bm{x} \in G} f(\bm{x}) \ket{\bm{x}} \mapsto \sum_{\bm{a} \in G} \cht_G(f)(\bm{a}) \ket{\bm{a}},
\end{equation}
where normalization of the input state is omitted for brevity. The function $\cht_G(f)$ may refer to either the additive form \eqref{eq:cht-add} or the multiplicative form \eqref{eq:cht-mult}. In the special case of a single basis state $\ket{a}$ in the group $\Z_N$, the quantum Hartley transform takes the form
\begin{equation}
	\label{eq:qht-N}
	\qht_N: \ket{a} \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big(\frac{2\pi a y}{N}\Big) \ket{y}.
\end{equation}

% Another concept we use is the \textbf{quantum Hartley transform} (QHT), which replaces complex exponentials with real sinusoidal kernels.

% \paragraph{Definition.}
% For $G = \mathbb{Z}_N$, define
% \[
% \qht_N: \ket{a} \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\left(\frac{2\pi a y}{N}\right) \ket{y}, \label{eq:qht_def}
% \]
% Where $\cas(x) = \cos(x) + \sin(x)$.

% For general abelian $G = \mathbb{Z}_{N_1} \oplus \cdots \oplus \mathbb{Z}_{N_k}$, the Hartley transform of $f : G \to \mathbb{R}$ is
% \begin{equation}
%     \qht(f)({\bm{a}}) = \frac{1}{\sqrt{|G|}} \sum_{{\bm{y}} \in G} \cas(2\pi \langle {\bm{a}}, \alpha({\bm{y}}) \rangle) f({\bm{y}}),
% \end{equation}
% where $\alpha({\bm{y}}) = y_1 / N_1 + \cdots + y_k / N_k$.

% \paragraph{Unitary Equivalence.}
% Using the identity
% \[
% \cas(2\pi \langle {\bm{a}}, \alpha({\bm{y}}) \rangle) = \frac{1 - i}{2}\chi({\bm{a}}, {\bm{y}}) + \frac{1 + i}{2}\chi(-{\bm{a}}, {\bm{y}}),
% \]
% we obtain:
% \begin{equation}
% \qht_G = \frac{1 - i}{2} \qft_G + \frac{1 + i}{2} \qft_G^*, \label{eq:QHT_QFT}
% \end{equation}
% showing that $\qht_G^2 = {1}$ and hence the $\qht$ is unitary.

% \paragraph{Quantum Formulation.}
% The quantum Hartley transform over $G$ maps:
% \[
% \sum_{{x} \in G} f({x}) \ket{{x}} \mapsto \sum_{{\bm{a}} \in G} \qht_G(f)({\bm{a}}) \ket{{\bm{a}}}.
% \]
% Because the cas function is real-valued, the resulting amplitudes remain real—a fact that plays a crucial role in real-amplitude quantum protocols.

@@

\begin{itemize}
    \item \textbf{Cycle Graph $C_n$:} The cycle graph is the Cayley graph of the cyclic group $\mathbb{Z}_n = \{0, 1, \dots, n-1\}$ with generating set $S = \{1, -1\}$. Each vertex $j \in \mathbb{Z}_n$ is connected to $j \pm 1 \mod n$.

    \item \textbf{Hypercube $Q_n$:} The $n$-dimensional Boolean hypercube is the Cayley graph of the group $(\mathbb{Z}_2)^n$ with generating set $S = \{e_1, \dots, e_n\}$, where $e_i$ flips the $i$-th bit. This graph is $n$-regular and bipartite.

@@

The generation algorithm $\mathsf{Gen}$ in Zhandry’s scheme works as follows (informally):
\begin{enumerate}
    \item On input $1^n$, sample a random group element $h \leftarrow G$. This $h$ will serve as the serial number $s$ for the banknote.
    \item Prepare two registers: the first is a group element register, initialized to $|0_G\rangle$ (an encoding of the identity element of $G$), and the second is an $X$-register initialized to $|x_0\rangle$.
    \item Apply the quantum Fourier transform over $G$ to the first register. This creates a uniform superposition over $G$: 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} |g\rangle \otimes |x_0\rangle\,.
    \] 
    \item Now apply the group action in a controlled manner: for each basis state $|g\rangle$ in the first register, apply $g$’s action on the second register. This transforms the joint state into 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} |g\rangle \otimes |g * x_0\rangle\,,
    \] 
    which is an entangled state between a superposition of group elements and their corresponding $X$-elements.
    \item Apply the inverse QFT (which is the adjoint of the QFT) on the first register. This recombines the amplitudes in the first register. The result (by properties of the QFT) can be shown to be 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} \omega^{\langle h,\,g \rangle} |g\rangle \otimes |g * x_0\rangle\,,
    \] 
    where $\omega^{\langle h,\,g \rangle}$ denotes the phase picked up that depends on $h$ and $g$. In fact, this state can be written as $|h\rangle \otimes \frac{1}{\sqrt{N}}\sum_{g \in G} |g * x_0\rangle$; essentially, the first register ends up in $|h\rangle$ (the serial number) and the second register is an equal superposition of all elements of $X$ in the orbit of $x_0$ (which is $X$ itself, since the action is transitive) but weighted by a phase related to $h$. This is the so-called group-action Fourier state.
    \item The first register (which contains $|h\rangle$) is measured in the computational basis, yielding the random serial number $g$ and collapsing the second register to a state that we will denote $|\$_h\rangle$. This $|\$_h\rangle$ is the actual quantum money state corresponding to serial $h$.
    \item Output the pair $(s = h,\; |\$\rangle = |\$_h\rangle)$.
\end{enumerate}

@@

\section{Hartley-Transform Money Scheme Construction}
Our scheme modifies the above Fourier-based construction by utilizing the \textbf{quantum Hartley transform (QHT)} in place of the QFT. Since the Hartley transform is closely related to the Fourier transform (essentially it produces the real and imaginary parts in a single real-valued transform), one might expect this substitution to be straightforward. And indeed, for the generation procedure it largely is: we will use the QHT to create real-amplitude superposition states.

Concretely, assume again a group action $*: G\times X \mapsto X$ with $|G|=N$. We fix the same generating set $S$ and basepoint $x_0$. The $\mathsf{Gen}_{H}$ (Hartley generation) algorithm:
\begin{enumerate}
    \item Sample a random $h \leftarrow G$ to be the serial number.
    \item Prepare $|0_G, x_0\rangle$ as before.
    \item Apply the \emph{quantum Hartley transform} over $G$ to the first register. The Hartley transform (on an abelian group of order $N$) is defined as:
    \[
       \mathrm{QHT}: |g\rangle \mapsto \frac{1}{\sqrt{N}} \sum_{k \in G} \cas \Big(\frac{2\pi \langle k,g \rangle}{N}\Big) |k\rangle,
    \] 
    where $\mathrm{cas}(\theta) = \cos \theta + \sin \theta$ and $\langle k,g \rangle$ denotes some pairing like an exponent or dot product (for cyclic groups this is just $kg$). In simpler terms, $\mathrm{QHT}(|0_G\rangle) = \frac{1}{\sqrt{N}}\sum_{k \in G} |k\rangle$ (because $\cas(0)=1$), \emph{the same uniform superposition as the QFT produces for $|0\rangle$}. So after this step, we again have $\frac{1}{\sqrt{N}}\sum_{g \in G} |g\rangle|x_0\rangle$, but crucially the coefficients are all \emph{real} and positive (in fact all $1/\sqrt{N}$).
    \item Perform the controlled group action by the first register on the second: $\frac{1}{\sqrt{N}}\sum_{g} |g\rangle |g*x_0\rangle$.
    \item Apply the inverse QHT on the first register (which being its own inverse up to a constant factor, is the same operation again since Hartley is symmetric like Fourier). Now here is where things differ: after this operation, the joint state will be 
    \[
       \frac{1}{\sqrt{N}} \sum_{g \in G} \cas\Big(\frac{2\pi \langle h,g \rangle}{N}\Big)\, |g\rangle \otimes |g*x_0\rangle.
    \] 
    Because the Hartley transform combines what would be cosine and sine terms, this state is not as nicely factorable as in the Fourier case. In the Fourier case, we got $|h\rangle \otimes \text{(orbit state)}$. In the Hartley case, we do not get a single basis state in the first register; instead we get a superposition in the first register with real coefficients $c_g = \cas(2\pi\langle h,g\rangle/N)$. However, note that $c_g$ is just some real number in $[-1,1]$ depending on $h$ and $g$. Another way to express the above state is:
    \[
      |\$\rangle = \frac{1}{\sqrt{N}} \sum_{x \in X} \Big(\frac{1}{\sqrt{N}}\sum_{g: g*x_0 = x} \cas\big(\frac{2\pi \langle h,g \rangle}{N}\big)\Big)\, |x\rangle.
    \] 
    Since for a given $x\in X$, all $g$ such that $g*x_0=x$ have the form $g = u h$ for a unique $u \in G$ (because the action is free and transitive), one can simplify the internal sum using properties of the cas function. But we need not dive into that here.
    \item No measurement of the first register is performed. In fact, we do not have a nice product state separation, so we cannot simply measure one register to get $h$. Instead, the output of generation is just the quantum state $|\$\rangle$ (which is entangled between what used to be two registers) and the known classical serial number $h$ (we know what $h$ we picked).
\end{enumerate}

@@

\section{Efficient Quantum Hartley Transform Implementation}
Before moving on to the verification algorithm, we briefly discuss how we implement the quantum Hartley transform (QHT) efficiently, as this is an important practical aspect of our scheme. The Fourier transform is well-known to have efficient circuits; for the Hartley transform, fewer results are available, so we contributed a new circuit construction.

The QHT we need is over an abelian group, which for concreteness one can think of as $\mathbb{Z}_N$ (the general finite abelian group case can usually be reduced to a direct sum of cyclic groups, applying QHT on each cyclic component). Classical Hartley transforms have a recursive structure similar to fast Fourier transforms. In fact, there are known formulas to express a Hartley transform of size $N$ in terms of two Hartley transforms of size $N/2$ plus some additional linear operations \cite{Bracewell1984}. We leveraged such structures to design a divide-and-conquer quantum circuit.

>>>>

\hypersetup{
	unicode = true,
    colorlinks = true,
    citecolor = blue,
    filecolor = blue,
    linkcolor = blue,
    urlcolor = blue,
    pdfstartview = {FitH},
}

\title{
  \textbf{Quantum Walks and Application to Quantum Money}\\[2em]
  \normalsize by\\[2em]
  \text{\large Ali Mousavi} \\[5em]
  \text{\shortstack{A Master's Thesis \\[0.5em]
   Submitted to the Department of Computing and Software\\[0.5em]
   McMaster University\\[0.5em]
   In Fulfillment of the Requirements\\[0.5em]
   For a Master's Degree}
  }

}

\date{April 2025}

@@

This thesis explores the foundations of quantum computation, focusing on quantum walks and their application to quantum money. Quantum walks, particularly continuous-time quantum walks based on group actions, serve as a powerful computational tool with applications in search algorithms and cryptographic protocols. We examine their mathematical structure and their advantages over classical random walks, emphasizing their efficiency in state evolution and probability distribution spreading. As a part of this work, we examine efficient implementations of transforms such as the Quantum Fourier Transform ($\qft$) and the Quantum Hartley Transform ($\qht$), analyzing their role in encoding quantum states for secure cryptographic applications. In particular, we discuss a novel instantiation of a quantum money scheme based on $\qht$, leveraging its unique properties for improved security and efficiency.
To ensure the robustness of this quantum money scheme, we develop a verification mechanism utilizing quantum walks. Unlike previous approaches, which rely on standard quantum state measurements, our method employs continuous-time quantum walks to authenticate quantum money, preventing counterfeiting while maintaining computational feasibility. Additionally, we present a detailed discussion on the efficient implementation of this scheme, including optimized circuit designs and error mitigation strategies.

@@

\medskip\noindent\textbf{This Work – Motivation and Contributions.} 
In this thesis, we adapt Zhandry’s group-action quantum money scheme by replacing its use of the Quantum Fourier Transform with the \emph{Quantum Hartley Transform} ($\qht$) over finite abelian groups. The motivation behind this substitution is multifold. First, using the Hartley transform causes the banknotes to have \emph{real} amplitudes instead of complex amplitudes. We believe this shift from complex to real quantum states may offer both computational and theoretical advantages. For example, certain mathematical identities hold for any two real orthonormal bases that do not generally hold for complex bases; such differences can affect the properties of quantum states and were a barrier in prior analyses (indeed, an attempted “quantum lightning” construction in \cite{Zhandry2022} failed due to properties of complex phases that do not arise with real amplitudes). Beyond this specific context, our work is a first step toward employing real-valued quantum transforms in quantum cryptography and quantum algorithms. To the best of our knowledge, this is the first significant quantum cryptographic construction that makes essential use of the quantum Hartley transform. We hope that exploring real-amplitude quantum states will inspire further research into new quantum algorithms and optimizations for real-valued transforms.

@@

Quantum computation is a model of information processing rooted in the principles of quantum mechanics. Unlike classical systems, which encode and manipulate information using deterministic bits, quantum systems store information in quantum states—vectors in complex Hilbert spaces—that can exhibit phenomena with no classical analogue. The most notable of these are \emph{superposition}, which allows quantum systems to be in multiple configurations simultaneously, and \emph{entanglement}, which enables correlations between parts of a system that cannot be explained classically. These characteristics endow quantum computers with computational abilities that can outperform classical counterparts in certain tasks. For a comprehensive introduction to quantum computation and quantum information we refer the reader to ~\cite{NielsenChuang2010, watrous2018theory, kaye2006introduction}.

@@

  \item \textbf{Composition of Systems.} If two quantum systems \( A \) and \( B \) are described by Hilbert spaces \( \mathcal{H}_A \) and \( \mathcal{H}_B \), respectively, then the combined system is described by the tensor product \( \mathcal{H}_A \otimes \mathcal{H}_B \). A product state of the form \( \ket{\psi_A} \otimes \ket{\psi_B} \) represents a system where \( A \) and \( B \) are uncorrelated. In general, however, the state may be entangled, and cannot be written as a product of subsystems.

  If unitary operators \( U_A \) and \( U_B \) act on \( \mathcal{H}_A \) and \( \mathcal{H}_B \), respectively, then the combined unitary \( U_A \otimes U_B \) acts on the joint system by applying \( U_A \) to subsystem \( A \) and \( U_B \) to subsystem \( B \), independently. That is,
  \[
  (U_A \otimes U_B)(\ket{\psi_A} \otimes \ket{\psi_B}) = (U_A \ket{\psi_A}) \otimes (U_B \ket{\psi_B}).
  \]
  This operation preserves the separability of product states, and more generally, acts linearly on the full Hilbert space \( \mathcal{H}_A \otimes \mathcal{H}_B \).

\end{enumerate}

@@

\section{The Quantum Fourier Transform}
The \textit{Quantum Fourier Transform $\qft$} is a quantum analogue of the classical discrete Fourier transform and plays a fundamental role in many quantum algorithms, including Shor’s factoring algorithm and quantum phase estimation. It enables the transformation of quantum states into the frequency domain, revealing periodicities and structure that may be hidden in the computational basis.

A common instance of the $\qft$ is over the cyclic group \( \mathbb{Z}_N  = \{0,1,...,N-1\}\). Given a state \( \ket{x} \in \mathbb{C}^N \), the $\qft$ over \( \mathbb{Z}_N \) maps it to
\[
\ket{x} \mapsto \frac{1}{\sqrt{N}} \sum_{k=0}^{N-1} e^{2\pi i xk / N} \ket{k}.
\]
When \( N = 2^n \), we can express the index \( x \) in binary as \( x = j_1 j_2 \ldots j_n \), where \( x = j_1 2^{n-1} + j_2 2^{n-2} + \cdots + j_n \). In this case, the $\qft$ has a beautiful product representation:
\[
\ket{j_1 \cdots j_n} \mapsto \frac{1}{2^{n/2}} \bigotimes_{k=1}^n \left( \ket{0} + e^{2\pi i 0.j_k j_{k+1} \cdots j_n} \ket{1} \right),
\]
where \( 0.j_k j_{k+1} \cdots j_n \) denotes the binary fraction \( \frac{j_k}{2} + \frac{j_{k+1}}{4} + \cdots + \frac{j_n}{2^{n-k+1}} \). This form reveals the tensor product structure of the $\qft$: the transformed state decomposes as a product of single-qubit superpositions, each involving successively finer phase rotations.

This structure allows for an efficient quantum circuit implementation. The circuit consists of a sequence of Hadamard gates and controlled phase gates \( R_k = \begin{bmatrix} 1 & 0 \\ 0 & e^{2\pi i / 2^k} \end{bmatrix} \), followed by a series of swap gates to reverse the order of qubits. For each qubit \( j_k \), a Hadamard gate is applied first, followed by controlled-\( R_2, R_3, \ldots \) gates from subsequent qubits. The number of gates required is \( \Theta(n^2) \), making the $\qft$ over \( 2^n \) dimensions efficiently implementable on a quantum computer \cite[Section 5.1]{NielsenChuang2010}.

In this work, we extend the notion of the $\qft$ to more general finite groups beyond \( \mathbb{Z}_N \). This generalization is essential for studying quantum walks on group-based graphs and analyzing quantum algorithms defined over non-abelian structures.

Let $G$ be an abelian group. The group of characters of $G$, denoted $\hat{G}$, consists of homomorphisms of the form $\chi(a, \cdot): G \to \mathbb{C}$ for each $a \in G$. If the group has the structure $G \cong \mathbb{Z}_{N_1} \oplus \cdots \oplus \mathbb{Z}_{N_k}$, then the character function $\chi(a, \cdot)$ can be expressed explicitly as
\[
\chi(a, x) = \omega_{N_1}^{a_1x_1} \cdots \omega_{N_k}^{a_kx_k}
\]
where $\omega_M = \exp(2\pi i / M)$ is a primitive $M$-th root of unity. Given a function $f: G \to \mathbb{C}$, its Fourier transform is defined as
\[
\hat{f}(a) = \frac{1}{\sqrt{\abs{G}}} \sum_{x \in G} \chi(a, x) f(x).
\]
Correspondingly, the quantum Fourier transform of a normalized state $\sum_{g \in G} f(g) \ket{g}$ is given by $\sum_{x \in G} \hat{f}(x) \ket{x}$. The $\qft$ over \( G \) is simply a tensor product of $\qft$s over each \( \mathbb{Z}_{N_i} \), and can therefore be implemented efficiently.

@@

Another concept we use is the \textbf{quantum Hartley transform} ($\qht$), which replaces complex exponentials with real sinusoidal kernels.
Let $N$ be a positive integer, and let $\Z_N$ denote the additive cyclic group of integers modulo $N$. The Hartley transform of a function $f: \Z_N \to \R$ is defined as the function $\cht_N(f): \Z_N \to \R$ given by
\[
\cht_N(f)(a) = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big(\frac{2 \pi ay}{N}\Big) f(y),
\]
where $\cas(x) = \cos(x) + \sin(x)$. Similar to the Fourier transform, $\cht_N$ is a linear operation, and one can verify that it is unitary. This definition extends naturally to general abelian groups. Suppose $G = \Z_{N_1} \oplus \Z_{N_2} \oplus \cdots \oplus \Z_{N_k}$ is a decomposition of an abelian group into cyclic components. For a function $f: G \to \R$, the Hartley transform $\cht_G(f): G \to \R$ is defined by
\begin{equation}
    \label{eq:cht-add}
    \cht_G(f)(a_1, \dots, a_k) = \frac{1}{\sqrt{\abs{G}}} \sum_{y_1 = 0}^{N_1 - 1} \cdots \sum_{y_k = 0}^{N_k - 1} \cas\Big(\frac{2 \pi a_1y_1}{N_1} + \cdots + \frac{2 \pi a_ky_k}{N_k}\Big) f(y_1, \dots, y_k),
\end{equation}
with $(a_1, \dots, a_k) \in G$. Introducing the notation $\alpha(\bm{y}) = y_1 / N_1 + \cdots + y_k / N_k$, a more concise form for $\cht_G$ is
\[
\cht_G(f)(\bm{a}) = \frac{1}{\sqrt{\abs{G}}} \sum_{\bm{y} \in G} \cas(2\pi \lrang{\bm{a}, \alpha(\bm{y})}) f(\bm{y}).
\]

@@

The quantum Hartley transform over an abelian group $G$ is given by
\begin{equation}
    \label{eq:qht}
    \qht_G: \sum_{\bm{x} \in G} f(\bm{x}) \ket{\bm{x}} \mapsto \sum_{\bm{a} \in G} \cht_G(f)(\bm{a}) \ket{\bm{a}},
\end{equation}
where normalization of the input state is omitted for brevity. The function $\cht_G(f)$ may refer to either the additive form \eqref{eq:cht-add} or the multiplicative form \eqref{eq:cht-mult}. In the special case of a single basis state $\ket{a}$ in the group $\Z_N$, the quantum Hartley transform takes the form
\begin{equation}
	\label{eq:qht-N}
	\qht_N: \ket{a} \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big(\frac{2\pi a y}{N}\Big) \ket{y}.
\end{equation}
The unitary operator $\qht_N$ can be implemented efficiently, as will be shown in Section~\ref{sec:efficient_qht}. When using the multiplicative form~\eqref{eq:cht-mult}, the unitary $\qht_G$ decomposes as a tensor product of unitaries of the form $\qht_N$.

@@

\begin{itemize}
    \item \textbf{Cycle Graph $C_n$:} The cycle graph is the Cayley graph of the cyclic group $\mathbb{Z}_n = \{0, 1, \dots, n-1\}$ with generating set $S = \{1, -1\}$. Each vertex $j \in \mathbb{Z}_n$ is connected to $j \pm 1 \mod n$. This graph is of particular interest in this work, as we will make use of it in chapter~\ref{chap:verification}.
    \begin{center}
        \includegraphics[width = 5cm]{cyclic_graph.pdf}
    \end{center}
    \item \textbf{Hypercube $Q_n$:} The $n$-dimensional Boolean hypercube is the Cayley graph of the group $(\mathbb{Z}_2)^n$ with generating set $S = \{e_1, \dots, e_n\}$, where $e_i$ flips the $i$-th bit. This graph is $n$-regular and bipartite.

@@

The generation algorithm $\mathsf{Gen}$ in Zhandry’s scheme works as follows (informally):
\begin{enumerate}
    \item On input $1^n$, sample a random group element $h \leftarrow G$. This $h$ will serve as the serial number $s$ for the banknote.
    \item Prepare two registers: the first is a group element register, initialized to $|0_G\rangle$ (an encoding of the identity element of $G$), and the second is an $X$-register initialized to $|x_0\rangle$.
    \item Apply the quantum Fourier transform over $G$ to the first register. This creates a uniform superposition over $G$: 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} |g\rangle \otimes |x_0\rangle\,.
    \] 
    \item Now apply the group action in a controlled manner: for each basis state $|g\rangle$ in the first register, apply $g$’s action on the second register. This transforms the joint state into 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} |g\rangle \otimes |g * x_0\rangle\,,
    \] 
    which is an entangled state between a superposition of group elements and their corresponding $X$-elements.
    \item Apply the inverse $\qft$ (which is the adjoint of the $\qft$) on the first register. This recombines the amplitudes in the first register. The result (by properties of the $\qft$) can be shown to be 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} \omega^{\langle h,\,g \rangle} |g\rangle \otimes |g * x_0\rangle\,,
    \] 
    where $\omega^{\langle h,\,g \rangle}$ denotes the phase picked up that depends on $h$ and $g$. In fact, this state can be written as $|h\rangle \otimes \frac{1}{\sqrt{N}}\sum_{g \in G} |g * x_0\rangle$; essentially, the first register ends up in $|h\rangle$ (the serial number) and the second register is an equal superposition of all elements of $X$ in the orbit of $x_0$ (which is $X$ itself, since the action is transitive) but weighted by a phase related to $h$. This is the so-called group-action Fourier state.
    \item The first register (which contains $|h\rangle$) is measured in the computational basis, yielding the random serial number $g$ and collapsing the second register to a state that we will denote $|\$_h\rangle$. This $|\$_h\rangle$ is the actual quantum money state corresponding to serial $h$.
    \item Output the pair $(s = h,\; |\$\rangle = |\$_h\rangle)$.
\end{enumerate}

@@

\section{Hartley-Transform Money Scheme Construction}

Our scheme modifies the above Fourier-based construction by utilizing the \textbf{quantum Hartley transform ($\qht$)} in place of the $\qft$. Since the Hartley transform is closely related to the Fourier transform (essentially it produces the real and imaginary parts in a single real-valued transform), one might expect this substitution to be straightforward. And indeed, for the generation procedure it largely is: we will use the $\qht$ to create real-amplitude superposition states.

Concretely, assume again a group action $*: G\times X \mapsto X$ with $|G|=N$. For concreteness, we assume that $G = \Z_N$; the genralization of the scheme to a general group $G$ is straightforward. We fix the same generating set $S$ and basepoint $x_0$. The $\mathsf{Gen}_{H}$ (Hartley generation) algorithm:
\begin{enumerate}
    \item Sample a random $h \leftarrow G$ to be the serial number.
    \item Prepare $|0_G, x_0\rangle$ as before.
    \item Apply the \emph{quantum Hartley transform} over $G$ to the first register, as we saw in \ref{eq:qht-N}:
    \[
       \qht: |g\rangle \mapsto \frac{1}{\sqrt{N}} \sum_{k \in \mathbb{Z}_N} \cas \Big(\frac{2\pi  kg }{N}\Big) |k\rangle,
    \] 
    where $\mathrm{cas}(\theta) = \cos \theta + \sin \theta$. In simpler terms, $\qht(|0_G\rangle) = \frac{1}{\sqrt{N}}\sum_{k \in G} |k\rangle$ (because $\cas(0)=1$), \emph{the same uniform superposition as the $\qft$ produces for $|0\rangle$}. So after this step, we again have $\frac{1}{\sqrt{N}}\sum_{g \in G} |g\rangle|x_0\rangle$, but crucially the coefficients are all \emph{real} and positive (in fact all $1/\sqrt{N}$).
    \item Perform the controlled group action by the first register on the second: $\frac{1}{\sqrt{N}}\sum_{g} |g\rangle |g*x_0\rangle$.
    \item Apply the inverse $\qht$ on the first register (which being its own inverse up to a constant factor, is the same operation again since Hartley is symmetric like Fourier). Now here is where things differ: after this operation, the joint state will be 
    \[
       \frac{1}{\sqrt{N}} \sum_{g \in \mathbb{Z}_N} \cas\Big(\frac{2\pi hg}{N}\Big)\, |g\rangle \otimes |g*x_0\rangle.
    \] 
    Because the Hartley transform combines what would be cosine and sine terms, this state is not as nicely factorable as in the Fourier case. In the Fourier case, we got $|h\rangle \otimes \text{(orbit state)}$. In the Hartley case, we do not get a single basis state in the first register; instead we get a superposition in the first register with real coefficients $c_g = \cas(2\pi hg/N)$. However, note that $c_g$ is just some real number in $[-1,1]$ depending on $h$ and $g$. Another way to express the above state is:
    \[
      |\$\rangle = \frac{1}{\sqrt{N}} \sum_{x \in X} \Big(\frac{1}{\sqrt{N}}\sum_{g: g*x_0 = x} \cas\big(\frac{2\pi  hg}{N}\big)\Big)\, |x\rangle.
    \] 
    Since for a given $x\in X$, all $g$ such that $g*x_0=x$ have the form $g = u h$ for a unique $u \in G$ (because the action is free and transitive), one can simplify the internal sum using properties of the cas function. But we need not dive into that here.
    \item No measurement of the first register is performed. In fact, we do not have a nice product state separation, so we cannot simply measure one register to get $h$. Instead, the output of generation is just the quantum state $|\$\rangle$ (which is entangled between what used to be two registers) and the known classical serial number $h$ (we know what $h$ we picked).
\end{enumerate}

@@

\section{Efficient Quantum Hartley Transform Implementation}\label{sec:efficient_qht}
Before moving on to the verification algorithm, we briefly discuss how we implement the quantum Hartley transform ($\qht$) efficiently, as this is an important practical aspect of our scheme. The Fourier transform is well-known to have efficient circuits; for the Hartley transform, fewer results are available, so we contributed a new circuit construction.

The $\qht$ we need is over an abelian group, which for concreteness one can think of as $\mathbb{Z}_N$ (the general finite abelian group case can usually be reduced to a direct sum of cyclic groups, applying $\qht$ on each cyclic component). Classical Hartley transforms have a recursive structure similar to fast Fourier transforms. In fact, there are known formulas to express a Hartley transform of size $N$ in terms of two Hartley transforms of size $N/2$ plus some additional linear operations \cite{Bracewell1984}. We leveraged such structures to design a divide-and-conquer quantum circuit.
